Gametic Selection, Meiotic Drive, Sex Ratio Bias, and Transitions Between Sex Determination Systems

### 2017-02-01 LaTeX/Turnover.tex

We find that neo-W alleles can invade for a large number of selective regimes. 
To clarify the parameter space under which $\lambda_{W',XY}>1$, we consider several special cases. 
Firstly, if the \textbf{A} locus is unlinked to the ancestral sex-determining region ($r=1/2$), a more closely linked neo-W ($R<1/2$) can always invade because $\left( \hat{p}^\male_Y-\hat{p}^\male_X \right)=0$ such that the second term in \eqref{eq:lambda_neoW} disappears and invasion depends on the sign of $(r-R)$. 
Indeed, invasion typically occurs when the neo-W is more closely linked to the selected locus than the ancestral sex-determining region, Figure \ref{fig:Combination_Centimorgans}.
Secondly, we can simplify cases where $R>r$ using the special case where $R=1/2$ and $r<1/2$. 
In table \ref{tab:specialcases} we give the conditions where invasion occurs where we further assume that haploid selection only occurs during one phase in one sex (e.g., during male meiosis only) and equal dominance, $h^\female=h^\male$. 
Where there is no gametic competition and meiotic drive in one sex only, an unlinked neo-W can invade as long as the same allele is favoured in male and female diploid selection ($s^\female s^\male>0$, see Figure \ref{fig:Combination_Centimorgans}B), which is 50\% of the parameter space. 
Where there is no meiotic drive and gametic competition occurs in one sex only, an unlinked neo-W can invade as long as the same allele is favoured in male and female diploid selection and there are sex differences in selection of one type ($-s^\female(s^\female-s^\male)>0$, see Figure \ref{fig:Combination_Centimorgans}C,D), which is 25\% of the parameter space. 
These special cases indicate that neo-W invasion can occur for a relatively large fraction of parameter space, even if $R>r$.

\begin{table}[ht]
\centering
\smallskip
\caption{Invasion conditions for unlinked neo-W ($R=1/2$, $r<1/2$) with one form of haploid selection only}
\begin{tabular}{l c }
\hline\hline
  Assumptions & neo-W spreads ($\lambda_{W',XY}>1$) if \\ [0.5ex] \hline
%  $r=1/2$, $R<1/2$ & Always \\ [0.5ex] \hline
%  $r<1/2$, $R=1/2$ &  \\ [0.5ex] \hline
  $h^\male=h^\female$, $t^\female=t^\male=\alpha^\female_{\Delta}=0$ & $s^\female s^\male>0$ \\
  $h^\male=h^\female$, $t^\female=t^\male=\alpha^\male_{\Delta}=0$ & $s^\female s^\male>0$ \\
  $h^\male=h^\female$, $t^\female=\alpha^\female_{\Delta}=\alpha^\male_{\Delta}=0$ & $-s^\female(s^\female-s^\male)>0$ \\
  $h^\male=h^\female$, $t^\male=\alpha^\female_{\Delta}=\alpha^\male_{\Delta}=0$ & $-s^\female(s^\female-s^\male)>0$ \\
  \hline \hline
  \label{tab:specialcases}
 \end{tabular}
\end{table}

>>>>

We find that neo-W alleles can invade for a large number of selective regimes. 
To clarify the parameter space under which $\lambda_{W',XY}>1$, we consider several special cases. 
Firstly, if the \textbf{A} locus is unlinked to the ancestral sex-determining region ($r=1/2$), a more closely linked neo-W ($R<1/2$) can always invade because $\left( \hat{p}^\male_Y-\hat{p}^\male_X \right)=0$ such that the second term in \eqref{eq:lambda_neoW} disappears and invasion depends on the sign of $(r-R)$. 
Indeed, invasion typically occurs when the neo-W is more closely linked to the selected locus than the ancestral sex-determining region, Figure \ref{fig:Combination_Centimorgans}.
Secondly, we can simplify cases where $R>r$ using the special case where $R=1/2$ and $r<1/2$. 
In table \ref{tab:specialcases} we give the conditions where invasion occurs where we further assume that haploid selection only occurs during one phase in one sex (e.g., during male meiosis only) and equal dominance, $h^\female=h^\male$. 
Where there is no gametic competition and meiotic drive in one sex only, an unlinked neo-W can invade as long as the same allele is favoured in male and female diploid selection ($s^\female s^\male>0$, see Figure \ref{fig:Combination_Centimorgans}B), which is 50\% of the parameter space. 
Where there is no meiotic drive and gametic competition occurs in one sex only, an unlinked neo-W can invade as long as the same allele is favoured in male and female diploid selection and there are sex differences in selection of one type ($s^\female(s^\male-s^\female)>0$, see Figure \ref{fig:Combination_Centimorgans}C,D), which is 25\% of the parameter space. 
These special cases indicate that neo-W invasion can occur for a relatively large fraction of parameter space, even if $R>r$.

\begin{table}[ht]
\centering
\smallskip
\caption{Invasion conditions for unlinked neo-W ($R=1/2$, $r<1/2$) with one form of haploid selection only}
\begin{tabular}{l c }
\hline\hline
  Assumptions & neo-W spreads ($\lambda_{W',XY}>1$) if \\ [0.5ex] \hline
%  $r=1/2$, $R<1/2$ & Always \\ [0.5ex] \hline
%  $r<1/2$, $R=1/2$ &  \\ [0.5ex] \hline
  $h^\male=h^\female$, $t^\female=t^\male=\alpha^\female_{\Delta}=0$ & $s^\female s^\male>0$ \\
  $h^\male=h^\female$, $t^\female=t^\male=\alpha^\male_{\Delta}=0$ & $s^\female s^\male>0$ \\
  $h^\male=h^\female$, $t^\female=\alpha^\female_{\Delta}=\alpha^\male_{\Delta}=0$ & $s^\female(s^\male-s^\female)>0$ \\
  $h^\male=h^\female$, $t^\male=\alpha^\female_{\Delta}=\alpha^\male_{\Delta}=0$ & $s^\male(s^\female-s^\male)>0$ \\
  \hline \hline
  \label{tab:specialcases}
 \end{tabular}
\end{table}
